---
title: "Internet Access, in relationship to GDP and College Enrollment"
author: ""
date: "Friday, March 6, 2015"
output: html_document
---
###Purpose of Study
The internet is often a tool of democratization and progress, which can foster a society which is more productive, healthy, and free. However, the effect is difficult to measure, because a society which is already productive and democratic is more likely to grant access to the internet. This study intends to show what relationship exists between per capita internet use and indicators of social and civic development, with a focus on controlling as much as possible for everything else which would influence that development. We start with simple 2-variable correlation, then essentially do everything possible to remove it through controls, and see what remains.

###Variables
So far, only GDP per Capita, Web Access Per capita, and College Enrollment(Gross % of college-age people) have been investigated. These are pulled from the WDI indicators ans labeled `gdp.cap`,`web.cap`, and `ter.cap`  respectively.

###Creation of Dataframe
The code used here is a bit of a brute force approach...
```{r, message=FALSE, echo=FALSE}
library(WDI)
library(dplyr)
library(ggplot2)
```
```{r, message=FALSE}
DF<-cbind(rename(WDI(indicator="IT.NET.USER.P2"), web.cap=IT.NET.USER.P2),
          rename(WDI(indicator="SE.SEC.ENRR"), sec.cap= SE.SEC.ENRR),
          rename(WDI(indicator="SE.TER.ENRR"), ter.cap= SE.TER.ENRR),
          rename(WDI(indicator="SE.XPD.TOTL.GB.ZS"), edubux= SE.XPD.TOTL.GB.ZS),
          rename(WDI(indicator="NY.GDP.PCAP.CD"), gdp.cap= NY.GDP.PCAP.CD),
          rename(WDI(indicator="SH.HIV.1524.FE.ZS"), fHIV= SH.HIV.1524.FE.ZS),
          rename(WDI(indicator="SH.HIV.1524.MA.ZS"), mHIV= SH.HIV.1524.MA.ZS),
          rename(WDI(indicator="SP.MTR.1519.ZS"),teenmom= SP.MTR.1519.ZS),
          rename(WDI(indicator="SH.DYN.AIDS.ZS"), tHIV=SH.DYN.AIDS.ZS),
          rename(WDI(indicator="SP.DYN.CONU.ZS"), contrac=SP.DYN.CONU.ZS),
          rename(WDI(indicator="EG.ELC.ACCS.ZS"),elec.cap=EG.ELC.ACCS.ZS),
          rename(WDI(indicator="SE.ENR.TERT.FM.ZS"),ter.gdr=SE.ENR.TERT.FM.ZS),
          rename(WDI(indicator="SL.EMP.INSV.FE.ZS"),wrk.fem=SL.EMP.INSV.FE.ZS))
attach(DF)
DF2<-as.data.frame(cbind(country,year,web.cap,elec.cap,gdp.cap,sec.cap,ter.cap,edubux,fHIV,mHIV,tHIV,teenmom,contrac, ter.gdr, wrk.fem))
detach(DF)
DF<-DF2
rm(DF2)
```
Some of those not mentioned before are HIV rates among men and women, money spent on education, and women in the workforce. At this point there is a DataFrame with the variables of interest. Unfortunately they are all coded as factors, so a similar brute force method is used to get them all coded numerically.

```{r, message=FALSE}
DF$web.cap<-as.numeric(as.character(DF$web.cap))
DF$sec.cap<-as.numeric(as.character(DF$sec.cap))
DF$ter.cap<-as.numeric(as.character(DF$ter.cap))
DF$edubux<-as.numeric(as.character(DF$edubux))
DF$gdp.cap<-as.numeric(as.character(DF$gdp.cap))
DF$year<-as.numeric(as.character(DF$year))
DF$fHIV<-as.numeric(as.character(DF$fHIV))
DF$mHIV<-as.numeric(as.character(DF$mHIV))
DF$teenmom<-as.numeric(as.character(DF$teenmom))
DF$tHIV<-as.numeric(as.character(DF$tHIV))
DF$contrac<-as.numeric(as.character(DF$contrac))
DF$elec.cap<-as.numeric(as.character(DF$elec.cap))
DF$ter.gdr<-as.numeric(as.character(DF$ter.gdr))
DF$wrk.fem<-as.numeric(as.character(DF$wrk.fem))
```
At this point we have variables of interest for all countries, as well as averages for all regions, for all years from 2005 to 2011. Because including regional averages would in a sense double-count the data from those countries, it's good to cut down the data to just country statistics. The data is helpfully indexed in a way where the regions are all up at the beginning rather than mixed alphabetically with the rest of the data, so this is actually just a simple command.

```{r}
DF.c<-DF[239:1736,]
```
At this point use the filter command from dplyr to investigate all sorts of data.

###Univariate Analysis
(There are no categorical variables of interest. The numeric variables could be cut into intervals to force a categorical study, but this results in information loss. In lieu of that, here are three numeric graphs/analyses.)
```{r, echo=FALSE}
DF.c.2011<-filter(DF.c, year==2011)
ggplot(data=DF.c.2011, aes(x=web.cap))+
  geom_histogram(color="blue",fill="lightblue",aes(y=..density..), binwidth=10)+
  geom_density(color="firebrick")+
  labs(title="Internet access, 2011",
       x="% Pop. With Internet access",
       y="number of countries"
      
)
ggplot(data=DF.c.2011, aes(x=ter.cap))+
  geom_histogram(color="red",fill="orange",aes(y=..density..), binwidth=10)+
  geom_density(color="firebrick")+
  labs(title="Tertiary Education, 2011",
       x="% College Age people in College",
       y="number of countries"
       
  )
ggplot(data=DF.c.2011, aes(x=gdp.cap))+
  geom_histogram(color="yellow",fill="green",aes(y=..density..), binwidth=10000)+
  geom_density(color="firebrick")+
  labs(title="GDP, 2011",
       x="GDP",
       y="number of countries"
       )
       
```



All are right-skewed, GDP heavily so. This may give result in problems with assumptions for inference down the line. 

###Bivariate Analysis
One interesting measure to look at is how many kids attend college, with respect to their country's level of internet access.

```{r, echo=FALSE}
DF.c.2011<-filter(DF.c, year==2011)
ggplot(data=DF.c.2011, aes(x=web.cap, y=ter.cap, ))+
  geom_point(color="firebrick")+
  labs(title="Tertiary Education by Internet Access, 2011",
       x="% Pop. With Internet access",
       y="% College age pop. in College"
       )
```



There is a clear relationship between the two variables, though not quite as strong as might be expected. Of course, both of these would be expected to share a strong relationship with wealth, and if the graph is colored by wealth,

###Multivariate Analysis

```{r, echo=FALSE}
ggplot(data=DF.c.2011, aes(x=web.cap, y=ter.cap, color=gdp.cap))+
  geom_point()+
  labs(title="Tertiary Education by Internet Access, 2011",
       x="% Pop. With Internet access",
       y="% College age pop. in College"
       )
```



We do see that the color, indicating GDP per capita,  gets brighter as both internet access and college attendance increase. This can be shown with scatterplots as well.
```{r, echo=FALSE}
ggplot(data=DF.c.2011, aes(x=web.cap, y=gdp.cap))+
  geom_point(colour="goldenrod")+
  labs(title="Web Access by GDP, 2011",
       x="% Pop. With Internet Access",
       y="GDP per capita"
       )
```
```{r, echo=FALSE}
ggplot(data=DF.c.2011, aes(x=ter.cap, y=gdp.cap,))+
  geom_point(colour="cyan")+
  labs(title="Tertiary Education by GDP, 2011",
       x="% College Age in College",
       y="GDP per capita"
       )
```



Clearly these relationships are a bit stronger.


