---
title: "Internet Access, in relationship to GDP and College Enrollment"
author: ""
date: "Friday, April 24, 2015"
output: html_document
---
###Purpose of Study
The internet is often a tool of democratization and progress, which can foster a society which is more productive, healthy, and free. However, the effect is difficult to measure, because a society which is already productive and democratic is more likely to grant access to the internet. This study intends to show what relationship exists between per capita internet use and indicators of social and civic development, with a focus on controlling as much as possible for everything else which would influence that development. We start with simple 2-variable correlation, then essentially do everything possible to remove it through controls, and see what remains.

###Variables
So far, only GDP per Capita, Web Access Per capita, and College Enrollment(Gross % of college-age people) have been investigated. These are pulled from the WDI indicators ans labeled `gdp.cap`,`web.cap`, and `ter.cap`  respectively.

###Creation of Dataframe
The code used here is a bit of a brute force approach...
```{r, message=FALSE, echo=FALSE}
library(WDI)
library(dplyr)
library(ggplot2)
```
```{r, message=FALSE}
DF<-cbind(rename(WDI(indicator="IT.NET.USER.P2"), web.cap=IT.NET.USER.P2),
          rename(WDI(indicator="SE.SEC.ENRR"), sec.cap= SE.SEC.ENRR),
          rename(WDI(indicator="SE.TER.ENRR"), ter.cap= SE.TER.ENRR),
          rename(WDI(indicator="SE.XPD.TOTL.GB.ZS"), edubux= SE.XPD.TOTL.GB.ZS),
          rename(WDI(indicator="NY.GDP.PCAP.CD"), gdp.cap= NY.GDP.PCAP.CD),
          rename(WDI(indicator="SH.HIV.1524.FE.ZS"), fHIV= SH.HIV.1524.FE.ZS),
          rename(WDI(indicator="SH.HIV.1524.MA.ZS"), mHIV= SH.HIV.1524.MA.ZS),
          rename(WDI(indicator="SP.MTR.1519.ZS"),teenmom= SP.MTR.1519.ZS),
          rename(WDI(indicator="SH.DYN.AIDS.ZS"), tHIV=SH.DYN.AIDS.ZS),
          rename(WDI(indicator="SP.DYN.CONU.ZS"), contrac=SP.DYN.CONU.ZS),
          rename(WDI(indicator="EG.ELC.ACCS.ZS"),elec.cap=EG.ELC.ACCS.ZS),
          rename(WDI(indicator="SE.ENR.TERT.FM.ZS"),ter.gdr=SE.ENR.TERT.FM.ZS),
          rename(WDI(indicator="SL.EMP.INSV.FE.ZS"),wrk.fem=SL.EMP.INSV.FE.ZS))
attach(DF)
DF2<-as.data.frame(cbind(country,year,web.cap,elec.cap,gdp.cap,sec.cap,ter.cap,edubux,fHIV,mHIV,tHIV,teenmom,contrac, ter.gdr, wrk.fem))
detach(DF)
DF<-DF2
rm(DF2)
```
Some of those not mentioned before are HIV rates among men and women, money spent on education, and women in the workforce. At this point there is a DataFrame with the variables of interest. Unfortunately they are all coded as factors, so a similar brute force method is used to get them all coded numerically.

```{r, message=FALSE}
DF$web.cap<-as.numeric(as.character(DF$web.cap))
DF$sec.cap<-as.numeric(as.character(DF$sec.cap))
DF$ter.cap<-as.numeric(as.character(DF$ter.cap))
DF$edubux<-as.numeric(as.character(DF$edubux))
DF$gdp.cap<-as.numeric(as.character(DF$gdp.cap))
DF$year<-as.numeric(as.character(DF$year))
DF$fHIV<-as.numeric(as.character(DF$fHIV))
DF$mHIV<-as.numeric(as.character(DF$mHIV))
DF$teenmom<-as.numeric(as.character(DF$teenmom))
DF$tHIV<-as.numeric(as.character(DF$tHIV))
DF$contrac<-as.numeric(as.character(DF$contrac))
DF$elec.cap<-as.numeric(as.character(DF$elec.cap))
DF$ter.gdr<-as.numeric(as.character(DF$ter.gdr))
DF$wrk.fem<-as.numeric(as.character(DF$wrk.fem))
```
At this point we have variables of interest for all countries, as well as averages for all regions, for all years from 2005 to 2011. Because including regional averages would in a sense double-count the data from those countries, it's good to cut down the data to just country statistics. The data is helpfully indexed in a way where the regions are all up at the beginning rather than mixed alphabetically with the rest of the data, so this is actually just a simple command.

```{r}
DF.c<-DF[239:1736,]
```
At this point use the filter command from dplyr to investigate all sorts of data.

###Univariate Analysis
(There are no categorical variables of interest. The numeric variables could be cut into intervals to force a categorical study, but this results in information loss. In lieu of that, here are three numeric graphs/analyses.)
```{r, echo=FALSE}
DF.c.2011<-filter(DF.c, year==2011)
ggplot(data=DF.c.2011, aes(x=web.cap))+
  geom_histogram(color="blue",fill="lightblue",aes(y=..density..), binwidth=10)+
  geom_density(color="firebrick")+
  labs(title="Internet access, 2011",
       x="% Pop. With Internet access",
       y="number of countries"
      
)
ggplot(data=DF.c.2011, aes(x=ter.cap))+
  geom_histogram(color="red",fill="orange",aes(y=..density..), binwidth=10)+
  geom_density(color="firebrick")+
  labs(title="Tertiary Education, 2011",
       x="% College Age people in College",
       y="number of countries"
       
  )
ggplot(data=DF.c.2011, aes(x=gdp.cap))+
  geom_histogram(color="yellow",fill="green",aes(y=..density..), binwidth=10000)+
  geom_density(color="firebrick")+
  labs(title="GDP, 2011",
       x="GDP",
       y="number of countries"
       )
       
```



All are right-skewed, GDP heavily so. This may give result in problems with assumptions for inference down the line. 

###Bivariate Analysis
One interesting measure to look at is how many kids attend college, with respect to their country's level of internet access.

```{r, echo=FALSE}
DF.c.2011<-filter(DF.c, year==2011)
ggplot(data=DF.c.2011, aes(x=web.cap, y=ter.cap, ))+
  geom_point(color="firebrick")+
  labs(title="Tertiary Education by Internet Access, 2011",
       x="% Pop. With Internet access",
       y="% College age pop. in College"
       )
```



There is a clear relationship between the two variables, though not quite as strong as might be expected. Of course, both of these would be expected to share a strong relationship with wealth, and if the graph is colored by wealth,

###Multivariate Analysis

```{r, echo=FALSE}
ggplot(data=DF.c.2011, aes(x=web.cap, y=ter.cap, color=gdp.cap))+
  geom_point()+
  labs(title="Tertiary Education by Internet Access, 2011",
       x="% Pop. With Internet access",
       y="% College age pop. in College"
       )
```



We do see that the color, indicating GDP per capita,  gets brighter as both internet access and college attendance increase. This can be shown with scatterplots as well.
```{r, echo=FALSE}
ggplot(data=DF.c.2011, aes(x=web.cap, y=gdp.cap))+
  geom_point(colour="goldenrod")+
  labs(title="Web Access by GDP, 2011",
       x="% Pop. With Internet Access",
       y="GDP per capita"
       )
```
```{r, echo=FALSE}
ggplot(data=DF.c.2011, aes(x=ter.cap, y=gdp.cap,))+
  geom_point(colour="cyan")+
  labs(title="Tertiary Education by GDP, 2011",
       x="% College Age in College",
       y="GDP per capita"
       )
```



Clearly these relationships are a bit stronger.


I can divide the web data into quartiles and test if there are significant differences between them with aov.

```{r}
quartile<- (as.integer(cut(DF.c.2011$web.cap, quantile(DF.c.2011$web.cap, probs=0:4/4, na.rm=T), include.lowest=TRUE)))
DF.c.2011.q<-cbind(DF.c.2011,quartile)
summary(aov(gdp.cap~quartile, data=DF.c.2011.q))
```

With an off-the-charts low p-value, there is clear evidence that there are differences in GDP between the quartiles.


###Chi-Square
We can also run a chi square test, dividing one of the variables into 2 levels. Say, "low internet use" and "high internet use", and test our data, assuming the null hypothesis that there is no relationship.


```{r, echo=FALSE, message=FALSE}
half.web<- (as.integer(cut(DF.c.2011$web.cap, quantile(DF.c.2011$web.cap, probs=0:2/2, na.rm=T), include.lowest=TRUE)))
quartile.gdp<- (as.integer(cut(DF.c.2011$gdp.cap, quantile(DF.c.2011$gdp.cap, probs=0:4/4, na.rm=T), include.lowest=TRUE)))
```

```{r,echo=FALSE}
T1 <- xtabs(~half.web + quartile.gdp, data = DF.c.2011)
T1
T2 <- prop.table(T1, 2)
T2
barplot(T2, xlab="GDP quartile", ylab="Internet use, dark=low, light=high")
```
```{r,echo=FALSE}
chisq.test(T1)
```
Again, there is a clear difference. If we run a chi square test on each pair of two (a total of 6 tests), at an alpha of .05/6=.008, we find that all are significant to reject the Null Hypothesis that there is no relationship.

```{r}
Ta <- xtabs(~half.web + quartile.gdp, data = DF.c.2011[DF.c.2011$quartile.gdp==1|DF.c.2011$quartile.gdp==2])
Tb <- xtabs(~half.web + quartile.gdp, data = DF.c.2011[DF.c.2011$quartile.gdp==1|DF.c.2011$quartile.gdp==3])
Tc <- xtabs(~half.web + quartile.gdp, data = DF.c.2011[DF.c.2011$quartile.gdp==1|DF.c.2011$quartile.gdp==4])
Td <- xtabs(~half.web + quartile.gdp, data = DF.c.2011[DF.c.2011$quartile.gdp==2|DF.c.2011$quartile.gdp==3])
Te <- xtabs(~half.web + quartile.gdp, data = DF.c.2011[DF.c.2011$quartile.gdp==2|DF.c.2011$quartile.gdp==4])
Tf <- xtabs(~half.web + quartile.gdp, data = DF.c.2011[DF.c.2011$quartile.gdp==3|DF.c.2011$quartile.gdp==4])

chisq.test(Ta)
chisq.test(Tb)
chisq.test(Tc)
chisq.test(Td)
chisq.test(Te)
chisq.test(Tf)
```

###Correlation
Let's look at the linear relationship between GDP and pursuit of college education.
```{r, echo=FALSE}
ggplot(data=DF.c.2011, aes(y=ter.cap, x=gdp.cap,))+
  geom_point(colour="purple")+
  geom_smooth(method=lm)+
  labs(title="Tertiary Education by GDP, 2011",
       y="% College age in College",
       x="% GDP per capita"
       )
```
Though there is a clear positive trend, this line doesn't seem to fit the data all too well. A transformation may be needed down the line, but for now, let's look at the correlation coefficient.

```{r}
cor(y=DF.c.2011$ter.cap, x=DF.c.2011$gdp.cap, use="complete")
```
This confirms the intuitive judgment. The association is positive, but weak.

Squaring this coefficient gives 0.2835543, so it can be said that GDP explains about 28.4% of the variance in pursuit of College per capita.




###Moderation
The possible moderator of interest is web access. To rephrase the purpose of study for this variable: Does access to internet change the existing relationship between GDP and pursuit of education?

```{r, echo=FALSE}
ggplot(data=DF.c.2011, aes(y=ter.cap, x=gdp.cap,color=(half.web==1)))+
  geom_point(colour=half.web)+
  geom_smooth(method=lm)+
  labs(title="Tertiary Education by GDP, 2011",
       y="% College age in College",
       x="% GDP per capita"
       )
```

It appears it does! We can test this by creating a model with an interaction term between GDP and internet access, and assuming the null hypothesis that such a term has a coefficient of zero. A low p value would cause us to reject this null and conclude that an interaction exists.

```{r}
summary(lm(ter.cap~ gdp.cap + half.web + gdp.cap*(half.web-1), data=DF.c.2011)) #The "-1" turns half.web into a dummy with levels 0 and 1
```
While it would seem intuitive that such an interaction would exist, the test has not found adequate evidence for it. The p-values of the terms are found under "Pr(>|t|)". If we are testing at alpha=.05, the interaction term is not significant.

###Logistic Regression
If the response of interest whether one's internet access was below or above the median, one could use per-capita GDP and the per-capita college attendance to predict this using Logistic Regression. This regression model is as follows:
$$ \hat{P}(Upper Half)= \frac{e^{\hat\beta_{0}+\hat\beta_{1}(GDP)+\hat\beta_{2}(College)}}
{1+e^{\hat\beta_{0}+\hat\beta_{1}(GDP)+\hat\beta_{2}(College)}} $$
The $\hat\beta$ terms are predicted coefficients of the variables. The interpretation is not very intuitive, but they are the per-unit increase in the log-odds of a country being in the top half in web access.

For each variable, a test is performed on the null hypothesis that it has no linear relationship to the log-odds, so
$$ H_{0}: \beta=0 $$
$$ H_{a}: \beta \neq 0 $$

Using the glm command in R:
```{r}
logit<-glm(data=DF.c.2011, (half.web-1)~ gdp.cap+ter.cap, family=binomial(link="logit"))
summary(logit)
```
The resulting p-values would suggest that GDP is certainly related to web access, but we can't  determine that per-capita college enrollment is, using this logit model. 